SMT-Boosted Security Types for Low-Level MPC

### 2025-11-24 toplas26/sty.tex
checkpoint

To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
intuitively meaning that all the share types in $\Gamma$ are sound with respect
to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
Property (i) ensures that views do not reveal information about secrets.
Property (ii) ensures the key property of sharing-- that individual shares
don't reveal more about secrets than their reconstruction does. Property
(iii) ensures that public values are independent of outputs or secrets. Properties
(iv) and (v) ensure that OTPs used for generating shares or for
secure openings, respectively, are included in $R$.
\begin{definition}
  \label{def-OK}
  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    $.
  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
  \item $\sep{\progd(\prog)}{(S \cup O)}{\setit{\phi_1,\ldots,\phi_j}}$
  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}
With the definition of ${OK}$ in place, we can define what we mean by soundness
of $\genshares$. In this definition, the first condition establishes that
initial shares are independent of honest secrets, and the second establishes
that they reveal no more than their reconstruction. These conditions are
obviously met by ``vanilla'' additive or Shamir shares which generate
shares in uniform marginal distributions.
\begin{definition}
  \label{def-genshares}
  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
  with $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  we have for all $a,b \in \setit{1,2}$:
  \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
    \end{enumerate}
\end{definition}

Next we show that, given well-formed typing environments, share type judgements preserve
the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
known semi-homomorphic properties of non-interactive operations that are syntactically
verified in typing judgements.
\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  and also $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$
  Then for all $a,b \in \setit{1,2}$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,\phi^b} \cup
      % \setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1, \phi_i^2), \phi_i^b}}$
    \end{enumerate}
\end{lemma}
\begin{proof}
  The proof is by induction on the derivation of $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$.
  In the base case where the judgement follows by the \TirName{Shared} rule, the result
  follows trivially by Definition \ref{def-OK}. In cases \TirName{HEAdd}, \TirName{HEAddPub}, and
  \TirName{HEMultPub} the result follows by Definition \ref{def-OK} and known
  semi-homorphic properties of secret sharing \cite{10.1007/978-3-030-68869-1_3}.
  Case \TirName{ShareEntails} follows by correctness of the SMT interpretation of
  protocols as in Theorem \ref{theorem-toeq}.
\end{proof}

>>>>

To proceed, we define a well-formedness property of tuples $\prog,\Gamma,R$
intuitively meaning that all the share types in $\Gamma$ are sound with respect
to $\prog$, and consumed randomness in $R$ as OTPs for initial sharing.
Property (i) ensures that views do not reveal information about secrets.
Property (ii) ensures the key property of sharing-- that individual shares
don't reveal more about secrets than their reconstruction does. Property
(iii) ensures that public values are independent of outputs or secrets. Properties
(iv) and (v) ensure that OTPs used for generating shares or for
secure openings, respectively, are included in $R$.
\begin{definition}
  \label{def-OK}
  Given $\prog,\Gamma,R$ where $\iov(\prog) = S \cup M \cup P \cup O$. For all $a,b\in \setit{1,2}$ with $a \ne b$ let:
  \begin{eqnarray*}
    %\Phi_a^m & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \\
    %\Phi_a^p & \defeq & \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma}\\
    \Phi_a & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
                        \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
    \Phi^p & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \pubty) \in \Gamma}\\
    \Phi^o & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \outty) \in \Gamma}
  \end{eqnarray*}
  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
  hold:
  \begin{enumerate}[\hspace{5mm}i.]
  \item For all $a,b \in \setit{1,2}$ we have
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \Phi^o}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o}
    $.
  \item For all $a,b \in \setit{1,2}$ and $([\phi_1,\phi_2] : \sharety) \in \Gamma$ we have:
    %$\condsep{\progd(\prog)}{\setit{\recon(\phi_i^1,\phi_i^2)}}{\setit{\phi_i^b}}{\recon(\phi_i^1,\phi_i^2)}$
    $$\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2)}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$$
  \item $\sep{\progd(\prog)}{(S \cup O)}{\Phi^p}$
  \item  For all $a,b \in \setit{1,2}$ and if
    $([x_1,x_2] : \sharety) \in \Gamma$ with $x_a \in \imesgs{M}{\setit{b}}{\setit{a}}$ then
    $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
    for some $\phi$ and $\rx{w}{\cid} \in R$.
  \item For all $\phi \in \Phi^p$ if $\toeq{\prog} \models \phi \eop \phi' + \rx{w}{\cid_{\oid}}$ for some $\phi'$
    and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
  \end{enumerate} 
\end{definition}
%\begin{definition}
%  \label{def-OK}
%  Given $\prog,\Gamma,R$ where $\secrets(\prog) = S$ and:
%  $$
%  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
%  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
%  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
%  $$
%  Then we write $\prog,\Gamma,R : {OK}$ iff all of the following conditions
%  hold:
%  \begin{enumerate}[\hspace{5mm}i.]
%  \item For all $a,b \in \setit{1,2}$ we have
%    $
%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi^o_1,\ldots,\phi^o_k}}
%    = 
%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
%      %\setit{\phi_1,\ldots,\phi_j}
%      \setit{\phi^o_1,\ldots,\phi^o_k}}
%    $.
%  \item For all $a,b \in \setit{1,2}$ and $0 < i \le n$ we have
%    %$\condsep{\progd(\prog)}{\setit{\recon(\phi_i^1,\phi_i^2)}}{\setit{\phi_i^b}}{\recon(\phi_i^1,\phi_i^2)}$
%    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2)}} =
%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(\phi_i^1,\phi_i^2), \phi_i^b}}$
%  \item $\sep{\progd(\prog)}{(S \cup O)}{\setit{\phi_1,\ldots,\phi_j}}$
%  \item  For all $a,b \in \setit{1,2}$ and $0 < i \le n$, if
%    $([x_1,x_2] : \sharety) \in \Gamma$ with $\genshares(\phi,\rx{w}{\cid}) = \phi_1,\phi_2$
%    and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$ 
%    for some $\phi$ and $\rx{w}{\cid}$ then $\rx{w}{\cid} \in R$.
%  \item For all $0 < i \le j$, if $\toeq{\prog} \models \phi_i \eop \phi + \rx{w}{\cid_{\oid}}$ for
%    some $\phi$  and $\rx{w}{\cid_{\oid}}$ then $\rx{w}{\cid_{\oid}} \in R$.
%  \end{enumerate} 
%\end{definition}
With the definition of ${OK}$ in place, we can define what we mean by soundness
of $\genshares$. In this definition, the first condition establishes that
initial shares are independent of honest secrets, and the second establishes
that they reveal no more than their reconstruction. These conditions are
obviously met by ``vanilla'' additive or Shamir shares which generate
shares in uniform marginal distributions.

\begin{definition}
  \label{def-genshares}
  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
  with $\secrets(\prog) = S$ and:
  $$
  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
  $$
  we have for all $a,b \in \setit{1,2}$:
  \begin{enumerate}[\hspace{5mm}i.]
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}
    =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
      %\setit{\phi_1,\ldots,\phi_j}
      \setit{\phi^o_1,\ldots,\phi^o_k}}$
    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
    \end{enumerate}
\end{definition}

%\begin{definition}
%  \label{def-genshares}
%  We say that $\genshares$ is \emph{sound} iff given any $\prog,\Gamma,R : {OK}$
%  and $\phi$ and $\rx{w}{\cid} \not \in R$ with $\genshares(\phi,\rx{w}{\cid}) =
%  \phi_1,\phi_2$ and $\toeq{\prog} \models x_1 \eop \phi_1 \wedge x_2 \eop \phi_2$
%  with $\secrets(\prog) = S$ and:
%  $$
%  \Gamma = \setit{[\phi_1^1,\phi_1^2] : \sharety, \ldots, [\phi_n^1,\phi_n^2] : \sharety}
%  \cup \setit{\pubx{\phi_1} : \pubty,\ldots,\pubx{\phi_j} : \pubty} \cup
%  \setit{\pubx{\phi^o_1} : \outty,\ldots,\pubx{\phi^o_k} : \outty}
%  $$
%  we have for all $a,b \in \setit{1,2}$:
%  \begin{enumerate}[\hspace{5mm}i.]
%    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b} \cup
%      %\setit{\phi_1,\ldots,\phi_j}
%      \setit{\phi^o_1,\ldots,\phi^o_k}}
%    =
%    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \setit{\phi_1^b,\ldots,\phi_n^b,x_b} \cup
%      %\setit{\phi_1,\ldots,\phi_j}
%      \setit{\phi^o_1,\ldots,\phi^o_k}}$
%    \item $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2)}} =
%      \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{\setit{\recon(x_1,x_2), x_b}}$
%    \end{enumerate}
%\end{definition}

Next we show that, given well-formed typing environments, share type judgements preserve
the important properties (i-ii) in Definitions \ref{def-OK}. This result follows from
known semi-homomorphic properties of non-interactive operations that are syntactically
verified in typing judgements.
\begin{lemma}
  \label{lemma-sty-noninteractive-sound}
  Given $\prog,\Gamma,R : {OK}$ where $\secrets(\prog) = S$. For all $a,b\in \setit{1,2}$ with $a \ne b$ let:
  \begin{eqnarray*}
    %\Phi_a^m & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \\
    %\Phi_a^p & \defeq & \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma}\\
    \Phi_a & \defeq & \setit{ \phi_a \mid ([\phi_1,\phi_2] : \sharety) \in \Gamma} \cup
                        \setit{ \phi_b \mid (\pubx{\recon(\phi_1,\phi_2)} : \sty) \in \Gamma} \\% \Phi_a^m \cup \Phi_a^p \\
    \Phi^o & \defeq & \setit{ \recon(\phi_1,\phi_2) \mid (\pubx{\recon(\phi_1,\phi_2)} : \outty) \in \Gamma}
  \end{eqnarray*} 
  and also suppose $\Gamma, \toeq{\prog} \vdash [\phi_1,\phi_2] : \sharety$.
  Then for all $a,b \in \setit{1,2}$ with $a \ne b$ the following conditions hold:
    \begin{enumerate}[\hspace{5mm}i.]
    \item
    $
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup \Phi^o}
    = 
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \setit{\phi_a} \cup \Phi^o}
    $
  \item 
    $\condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}} \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2)}} =
    \condd{\progtt(\prog)}{S_{\setit{\cid_a}}}{S_{\setit{\cid_b}}  \cup  \Phi_a \cup \Phi^o \cup \setit{\recon(\phi_1,\phi_2), \phi_b}}$
    \end{enumerate}
\end{lemma}
\begin{proof}
  The proof is by induction on the derivation of $\Gamma, \toeq{\prog} \vdash [\phi^1,\phi^2] : \sharety$.
  In the base case where the judgement follows by the \TirName{Shared} rule, the result
  follows trivially by Definition \ref{def-OK}. In cases \TirName{HEAdd}, \TirName{HEAddPub}, and
  \TirName{HEMultPub} the result follows by Definition \ref{def-OK} and known
  semi-homorphic properties of secret sharing \cite{10.1007/978-3-030-68869-1_3}.
  Case \TirName{ShareEntails} follows by correctness of the SMT interpretation of
  protocols as in Theorem \ref{theorem-toeq}.
\end{proof}
